### 2019-09-09 numerical_experiments.tex

\section{Numerical Experiments}\label{sec:numerical_experiments}
This section analyzes the performance of \methodAcronymROMs\ leveraging \spatialAcronym\ trial spaces on two benchmark problems: the Sod shock tube and compressible flow in a cavity. In each experiment, \methodAcronym\ is compared to the Galerkin and LSPG ROMs. Additionally, we assess the impact of the window size, time-step size, and time scheme on \methodAcronym. It is noted that both experiments are designed to test the \textit{reproductive} ability of the ROMs. Future state prediction and/or interpolation to new flow conditions is not considered as these problems introduce confounding factors (e.g., accuracy of the basis).

\subsection{Sod Shock Tube}
We first consider reduced-order models of the Sod shock tube problem. The system is governed by the compressible Euler equations in one dimension, 
\begin{equation}\label{eq:euler_1D}
    \frac{\partial \boldsymbol u}{\partial t} + \frac{\partial \boldsymbol F}{\partial \mathsf{x}} = 0, \quad
    \boldsymbol u= 
    \begin{Bmatrix} \rho \\ \rho u \\ \rho E \end{Bmatrix}, \quad 
    \boldsymbol F = \begin{Bmatrix} \rho u \\ \rho u^2 + p \\  u(\rho E + p) \end{Bmatrix},
\end{equation}
where $\boldsymbol u : \Omega \times [0,T] \rightarrow \RR{3}$ comprise the density, x-momentum, and total energy, $\mathsf{x} \in \Omega \defeq  [0,1]$ is the spatial domain, and 
$T = 1$ the final time. 
The problem setup is given by the initial conditions,

\begin{equation*}
\rho = 
\begin{cases} 
      1 & \mathsf{x}\leq 0.5 \\
      0.125 & \mathsf{x} > 0.5 
   \end{cases},
\qquad
p = 
\begin{cases} 
      1 & \mathsf{x}\leq 0.5 \\
      0.1 & \mathsf{x} > 0.5 
   \end{cases},
\qquad
u = 
\begin{cases} 
      0 & \mathsf{x}\leq 0.5 \\
      0 & \mathsf{x} > 0.5 
   \end{cases},
\end{equation*}
along with reflecting boundary conditions at $\mathsf{x}=0$ and $\mathsf{x}=1$.

\subsubsection{Description of FOM and Generation of Trial Space}
The 1D compressible Euler equations are solved using a finite volume method. The domain is partitioned into 500 cells of uniform width. The finite volume method uses the Rusanov flux~\cite{rusanov} at the cell interfaces. The FOM uses the Crank--Nicolson (CN) scheme, which is a linear multistep method defined by the coefficients $\alpha_0 = 1,\alpha_1 = -1, \beta_0 = \beta_1 = 1/2$, for temporal integration. The FOM is evolved for $t \in [0.0,1.0]$ at a time-step of $\Delta t = 0.002$. The \spatialAcronym\ trial space is constructed by executing Algorithm~\ref{alg:pod} with inputs $N_{\text{skip}}=2, \stateIntercept = \bz, K=46 $. The resulting trial space comprises $99.99\%$ of the snapshot matrix energy.
%  The solution is saved every other time-step, resulting in 500 solutions snapshots. These snapshots are used to generate a trial basis for the reduced-order models through POD. The trial basis is selected to be of dimension $K = 46$, which corresponds to a $99.99\%$ energy criterion. No affine transformation is used in the construction of the trial space, i.e., $\stateIntercept = \bz$.

\subsubsection{Description of Reduced-Order Models}
Reduced-order models based on the Galerkin, LSPG, and \methodAcronym\ approaches are considered. No hyper-reduction is considered 
in this example, i.e., $\stweightingMatArg{} = \mathbf{I}$. Details on the implementation of the 
different reduced-order models is as follows:
\begin{itemize}
\item \textbf{Galerkin ROM}: The Galerkin ROM is obtained through Galerkin projection of the FOM. The Galerkin ROM is evolved in time with the CN time scheme at 
a constant time-step size of $ \Delta t = 0.002$.

\item \textbf{LSPG ROM:} The LSPG ROM is built on top of the FOM discretization leveraging the CN time-scheme as previously described. Unless noted otherwise, the LSPG ROM uses a constant time-step size of $\Delta t = 0.002$. The nonlinear least-squares problem arising at each time-step instance is solved via the Gauss--Newton method. All Jacobians are computed via automatic differentiation~\cite{adolc}. The Gauss--Newton algorithm is deemed converged when the gradient norm is lower than $1e-4$. 
\item \textbf{\methodAcronymROM:} \methodAcronymROMs\ are solved via the direct and indirect methods with two different time-stepping schemes: 
\begin{itemize}
\item Direct Method: \methodAcronymROMs\ solved via the direct method are implemented for both the same CN discretization employed in the FOM and LSPG, as well as for the second-order explicit Adams Bashforth (AB2) discretization using a constant time-step size of $\Delta t = 0.0005$. The nonlinear least-squares problem 
arising over each window is solved via the Gauss--Newton method. All Jacobians are again computed via automatic differentiation, and the Gauss--Newton algorithm is deemed converged when the gradient norm is lower than $1e-4$. Critically, it is noted that the Jacobian 
matrix over a window is constructed by assembling local (dense) Jacobians. The global Jacobian over a single window is then stored in a compressed sparse row format. Uniform quadrature weights are used for evaluating the integral in~\eqref{eq:obj_gen_slab}.

\item Indirect Method: Two \methodAcronymROMs\ solved via the indirect method are implemented. The first uses the same CN discretization at at time-step size of $\Delta t = 0.002$, while the second uses the AB2 discretization using a time-step size of $\Delta t = 0.0005$. The coupled two-point boundary 
problem is solved via the Forward Backward Sweep Method. The action of the Jacobian transpose on vectors is computed via automatic differentiation. 
\end{itemize}
\end{itemize}

%We explore the evaluation of the \methodAcronym\ ROM using both indirect and direct solution methodologies. 
%\subsubsection{Indirect Method}
%The indirect \methodAcronym-ROM solves the coupled boundary value problem using the Forward-Backward Sweep Method (FBSM) described in Section~\ref{sec:fbsm}. A secondd-order collocation method with a time-step size of $\Delta t = 0.002$ and with collocation points placed at the Lobatto points is used for discretization of both the forward and backwards problems. The action of the Jacobian transpose on the residual and adjoint state required for the backwards problem is computed through automatic differentiation~\cite{adolc,pyadolc}. 
%
%\subsubsection{Direct Method}
%The direct \methodAcronym-ROM uses the same second-order collocation method with uniformly distributed collocation points for temporal discretization of both the state 
%and integrand. The resulting least-squares problem is solved using the Gauss-Newton method. The inner Gauss-Newton iterations are solved using the conjugate gradient 
%method preconditioned by the inverse approximate Hessian about the initial guess.

\subsubsection{Numerical Results}
The impact of the window size on \methodAcronymROMs\ is first assessed. A set of \methodAcronymROMs\ that minimize the residual over windows of constant size 
$\DeltaSlabArg{n} \equiv \DeltaSlabArg{} =  .002$, $0.004$ ,$0.008$, $0.02$, $0.04$, $0.10$, $0.20$, $1.0$ are considered. The standard Galerkin and LSPG ROMs are additionally considered. Results are shown for the direct method with the CN time-discretization scheme; a comparison of different time-marching methods and direct/indirect solution techniques will be provided later in this section. 
First, Figure~\ref{fig:sod_density} presents the density solutions predicted by the various ROMs at $t = \{0.5,1.0\}$. Figure~\ref{fig:sod_xt} shows 
$x-t$ diagrams for the same density solutions. From Figures~\ref{fig:sod_density} and~\ref{fig:sod_xt}, it is observed that the LSPG and Galerkin ROMs accurately characterize 
the system: they correctly track the shock location, expansion waves, etc. Both predictions, however, are observed to by highly oscillatory. These oscillations are 
not physical and can lead to numerical blowup; e.g., due to negative pressure. The \methodAcronymROMs\ produce less oscillatory solutions than both 
the Galerkin and LSPG ROMs. Critically, it is seen that, as the window size over which the residual is minimized grows, the solution becomes less oscillatory. 
When the residual is minimized over the entire space--time domain, the solution displays no oscillations.

%The impact of the window-size over which the residual is minimized on the ROM solution is first examined. To this end, reduced-order models that minimize the residual over windows of size $\Delta \tau = \Delta t \times [2,5,10,20,50,100,200,1000,2000]$ are considered. Figure~\ref{fig:xtrho} summarizes the performance of the various reduced-order models. Figure~\ref{fig:sod_error_a} shows the integrated $L^2$ state error relative to the Galerkin ROM. Figure~\ref{fig:sod_error_b} shows the space-time residual of the various reduced-order models as measured realitive to the full-order model. In Figure~\ref{fig:sod_error_a}, it is seen that the optimal ROM as measured by the integrated $L^2$ state error occurs at an intermediate window size. This result is perhaps somewhat surprising as one may expect the optimal $L^2$ error to occur when the window size spans the entire temporal domain, in which case the total space-time residual is minimized. As is evident from Figure~\ref{fig:sod_error_a}, however, a lower space-time residual does not necesarily imply a lower error. Figure~\ref{fig:sod_error_b} shows the total integrated space-time residual of the various ROMs as measured with respect to the projected full-order model solution. Increasing the window-size over which the residual is minimized leads to a lower over-all space-time residual. Further, when the residual is minimized over the entire space-time domain ($\Delta \tau =2000$), it is lower than that of the projected full-order model. This numerical evidence supports the theoretical error analysis presented in Section~\ref{sec:error}. In this example, the global space-time residual is lower than that of the projected full-order model for $\Delta \tau > 0.01$.  
%
%Figure~\ref{fig:sod_density} shows density profiles predicted by the CWST-LSPG ROM at two different time instances along with integrated total variation of the solution as measured with respect to the full-order model. It is observed that, as the window size over which the residual is being minimized grows, the solution becomes more smooth. This is measured quantitatively by the total variation of the solution, which decreases monotonically as the window size grows. Next, Figure~\ref{fig:xtrho} shows $x-t$ diagrams for the density variable for the various reduced-order models. It is again seen that, as $\Delta \tau$ grows, the solution becomes more smooth. The increased smoothness comes at the tradeoff that the shock profiles become more smeared. 
%
%Lastly, Figure~\ref{fig:sod_walltime} shows the wall-time of the CWST-LSPG ROM as a function of window-size. The cost is measured with respect to the Galerkin ROM. Two main observations are made. First, the cost of the CWST-LSPG ROM increases as the window size grows. This increased cost is because, as the window size grows, more iterations of the forward and backwards problems are required to converge the solution. While a more sophisticated solution methodology may help mitigate this increased cost, the trends are expected to say the same. The second observation is that the CWST-LSPG ROM is significantly more expensive than the standard Galerkin ROM. As discussed in Section~\ref{sec:implement}, this increased cost is due to the need to iteratively solve the two-point boundary value problem. 
\begin{figure}
\begin{center}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/rho_t05_compare.pdf}
\caption{$t=0.5$}
\end{subfigure}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/rho_t1_compare.pdf}
\caption{$t=1$}
\end{subfigure}
\caption{Density profiles at various time instances.}
\label{fig:sod_density}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\begin{subfigure}[t]{0.48\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/xt_fom.pdf}
\caption{Full-order model}
\end{subfigure}
\begin{subfigure}[t]{0.48\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/xt_grom.pdf}
\caption{G-ROM}
\end{subfigure}
\begin{subfigure}[t]{0.48\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/xt_lspg.pdf}
\caption{LSPG}
\end{subfigure}
\begin{subfigure}[t]{0.48\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/xt_w1.pdf}
\caption{\methodAcronym: $\DeltaSlabArg{}= 0.002$}
\end{subfigure}
\begin{subfigure}[t]{0.48\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/xt_w50.pdf}
\caption{\methodAcronym: $\DeltaSlabArg{} = 0.1$}
\end{subfigure}
\begin{subfigure}[t]{0.48\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/xt_w500.pdf}
\caption{\methodAcronym: $\DeltaSlabArg{} = 1.0$}
\end{subfigure}
\caption{$x-t$ diagrams for the density fields as predicted by the FOM, G-ROM, LSPG-ROM, and \methodAcronymROMs.}
\label{fig:sod_xt}
\end{center}
\end{figure}

Next, Figure~\ref{fig:sod_error} shows integrated state errors and the objective function (i.e., the residual norm) over $t \in [0,1]$ for the various  ROMs. Most notably, it is observed 
that increasing the window size over which the residual is minimized does \textit{not} lead to a monotonic decrease in the integrated $\elltwo$  
error. Increasing the window size does, however, lead to a monotonic decrease in the total integrated residual. It is additionally noted that, although the integrated
$\elltwo$ error of 
the projected FOM solution is significantly lower than that of the various ROM solutions, the integrated residual norm of the projected FOM solution is \textit{higher} 
than all ROMs. Thus, although the projected FOM solution is more accurate in the $\elltwo$ norm, it does not satisfy the governing equations as well the ROM solutions.

\begin{figure}
\begin{center}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/error_vs_window_compare.pdf}
\caption{Integrated $\elltwo$ error of various ROMs.}
\label{fig:sod_error_a}
\end{subfigure}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/objective_vs_window_compare.pdf}
\caption{Integrated residual norm of various ROMs.} 
\label{fig:sod_error_b}
\end{subfigure}
\caption{Integrated performance metrics of the Galerkin, LSPG, and \methodAcronymROMs.  Note that the Galerkin and LSPG ROMs do not depend on the window size.} 
\label{fig:sod_error}
\end{center}
\end{figure}

We now examine the comparative performance of the direct and indirect solution techniques for the \methodAcronymROMs\ using various time discretization techniques. 
Figure~\ref{fig:sod_error_methods} 
shows the same integrated $\elltwo$ state errors and residual norms as in Figure~\ref{fig:sod_error}, but this time results are shown for 
the various \methodAcronymROMs. In both Figures~\ref{fig:sod_error_methods_a} and~\ref{fig:sod_error_methods_b}, it is observed that, although some 
deviations are observed, the \methodAcronym\ method is relatively insensitive to the solution technique (direct vs indirect) and underlying discretization scheme.
In particular, ROMs using the second order explicit AB2 scheme with a time-step size of $\Delta t = 0.0005$ provide similar results to the 
ROMs using the second-order CN scheme at a time-step size of $\Delta t = 0.002$. All methods display similar dependence on the window size: the optimal integrated 
$\elltwo$ state error occurs when the window size is $\DeltaSlabArg{} = 0.1$ and the residual decreases monotonically as the window size grows. 
\begin{figure}
\begin{center}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/error_vs_window.pdf}
\caption{Integrated $\elltwo$ error of the \methodAcronymROMs.}
\label{fig:sod_error_methods_a}
\end{subfigure}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/objective_vs_window.pdf}
\caption{Integrated residual norm of the \methodAcronymROMs.} 
\label{fig:sod_error_methods_b}
\end{subfigure}
\caption{Integrated performance metrics of various \methodAcronymROMs.} 
\label{fig:sod_error_methods}
\end{center}
\end{figure}

Next, Figure~\ref{fig:sod_walltime} provides the CPU wall-clock times for the various \methodAcronymROMs. The computational cost of all methods is seen to grow 
as the window size is increased. For indirect methods, this increase in cost is due to the fact that, as the window size 
grows, more iterations of the FBSM are required for convergence. For direct methods, the increase in cost is due to (1.) the cost 
associated with forming and solving the normal equations at each Gauss--Newton iteration and (2.) the increased number of 
Gauss--Newton iterations required for convergence. The cost of the FBSM method is observed to increase more rapidly 
than the direct methods. Encouragingly, \methodAcronymROMs\ based on the CN discretization
minimizing the full space--time residual (comprising 
500 time instances) only  
cost approximately one order of magnitude more than the case where the residual is minimized over a single time-step (i.e., LSPG). Also of interest is the 
fact that the direct method utilizing the AB2 time-discretization scheme costs only slightly more per a given window size than the direct method 
utilizing the CN time discretization. This is despite the fact that AB2 is evolved at a time step of $\Delta t = 0.0005$, while CN uses $\Delta t = 0.002$; thus 
AB2 contains 4 times more temporal degrees of freedom than CN.  Finally, it is emphasized that the 
results presented here are presented for standard algorithms (e.g., Guass--Newton and the FBSM). As mentioned in 
Remarks~\ref{remark:gaussnewton} and~\ref{remark:fbsm}, it is expected the computational cost 
of both the indirect methods and direct methods can be decreased through the use and/or development of 
algorithms tailored to the minimization problem.

\begin{figure}
\begin{center}
\begin{subfigure}[t]{0.45\textwidth}
%\includegraphics[width=1.\linewidth]{/Users/ejparis/PyROM/pyrom/sod/postProcess_wv/error_vs_window.pdf}
\includegraphics[width=1.\linewidth]{figs/sod/walltime_vs_window.pdf}
\label{fig:sod_error_a}
\end{subfigure}
\caption{Comparison of wall-clock times of the direct and indirect \methodAcronymROMs\ as a function of window size.} 
\label{fig:sod_walltime}
\end{center}
\end{figure}

Finally, we study the impact of the time-step size on the \methodAcronymROM\ results. We examine \methodAcronymROMs\ that use a
window size of $\Delta T = 0.1$, with time-step sizes of $\Delta t =$  $0.001$, $0.002$, $0.005$, $0.01$, $0.05$, $0.1$ (i.e.,  
$100$, $50$, $20$, $10$, $2$, and $1$ time-steps per window). We additionally 
consider LSPG ROMs leveraging the same set of time steps. Figure~\ref{fig:time_step_study} shows the integrated $\elltwo$ error 
of the various ROMs as well as the integrated residual norm. We observe that both the $\elltwo$ state error and integrated residual 
of the \methodAcronymROMs\ decrease and converge as the time-step size decreases. This is in direct contrast to the LSPG ROMs, in where both the $\elltwo$ 
error and integrated residual norm increase as the time-step size decreases. This result 
demonstrates that \methodAcronym\ overcomes the time-step dependence inherit to the LSPG approach. Lastly, Figure~\ref{fig:walltime_dtvar}
shows the relative wall-clock times of the \methodAcronymROMs\ with respect to the LSPG ROMs. It is seen that, for all time-step sizes considered, the \methodAcronym-ROMs 
are less than 6x the cost LSPG; despite the window sizes comprising up to 100 time steps. 
\begin{figure}
\begin{center}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/error_vs_window_dtvar.pdf}
\caption{Integrated $\elltwo$ error of the \methodAcronymROM.}
\label{fig:sod_error_methods_a}
\end{subfigure}
\begin{subfigure}[t]{0.45\textwidth}
\includegraphics[width=1.\linewidth]{figs/sod/objective_vs_window_dtvar.pdf}
\caption{Integrated residual norm of the \methodAcronymROM.} 
\label{fig:sod_error_methods_b}
\end{subfigure}
\caption{Integrated performance metrics of the Galerkin, LSPG, and \methodAcronymROMs} 
\label{fig:time_step_study}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\begin{subfigure}[t]{0.45\textwidth}
%\includegraphics[width=1.\linewidth]{/Users/ejparis/PyROM/pyrom/sod/postProcess_wv/error_vs_window.pdf}
\includegraphics[width=1.\linewidth]{figs/sod/walltime_vs_window_dtvar.pdf}
\label{fig:sod_error_a}
\end{subfigure}
\caption{Comparison of wall-clock times of the \methodAcronymROMs\ to LSPG ROMs as a function of time-step size.} 
\label{fig:walltime_dtvar}
\end{center}
\end{figure}

\subsubsection{Summary of Numerical Results for the Sod Shock Tube}
The key observations from the results of the first numerical example are: 
\begin{enumerate}
\item Increasing the window size over which the residual is minimized lead to more physically relevant solutions. Specifically, we observed that as the window size over which the residual was minimized grew, \methodAcronym\ lead to less oscillatory solutions.
\item Increasing the window size over which the residual is minimized does not necessary lead to a more accurate solution in the $\elltwo$ sense. We observed that minimizing the residual over an intermediary window size led to the most accurate results as measured by the $\elltwo$ norm.
\item The \methodAcronym\ displays time-step convergence: both the integrated $\elltwo$ state error and residual norm decreased and displayed time-step convergence as the time-step size decreased. This is in contrast to LSPG.
\item For all examples considered, in where the windows comprised up to 2000 time-steps, \methodAcronym\ with the direct method is between 1x and 10x the cost of the LSPG. The 
direct method was observed to be, in general, slightly more efficient and robust than the indirect method.
\end{enumerate}

\input{cavity}