### 2019-10-14 section3/tclsrm_formulation.tex

\input{section3/general_form}
\input{section3/spatial_projection}
\input{section3/space_time}

\subsection{\methodAcronym\ summary}
This section outlined the \methodAcronym\ method for model reduction. In
summary, \methodAcronym\ sequentially minimizes the time continuous FOM ODE
residual within the range of a space--time trial over time
windows subject to boundary conditions associated with continuity across time
windoes; i.e.,  
%\methodAcronym\  
%computes the generalized coordinates
%$\genstateArgnt{n}$, for $n = 1,\ldots,\nslabs$, as the solution to the
%minimization problems, 
\methodAcronym\ sequentially computes approximate solutions
$\approxstateArgnt{n}\in\stspace^n$, $n=1,\ldots,\nslabs$ that comprise
solutions to the optimization problem \eqref{eq:tclsrm}.
\KTC{I removed repeated equations here and just referred to their previous
definitions. This will cut down on the text.}
%\begin{equation*}
%\underset{\stateyArgnt{} \in \stspace^n}{\text{minimize } }
%\mathcal{J}^n(\stateyArgnt{}),
%\end{equation*}
%where
%\begin{align*}
%\mathcal{J}^n &\vcentcolon \statey \mapsto
%\frac{1}{2} \int_{\timeStartArg{n}}^{\timeEndArg{n}} \big[ \dot{\statey}(t)
%- \velocity(\stateyArg{}{t},t) \big]^T \stweightingMatArg{n} \big[
%\dot{\statey}(t) - \velocity(\stateyArg{}{t},t) \big] d t \\
%&\vcentcolon \RR{\fomdim}\otimes \timeSpaceArg{n} \rightarrow
%\RRplus,
%\end{align*}
%subject to the boundary conditions.

For \spatialAcronym\ trial subspaces, the stationary conditions for the
residual minimization problem can be derived via the Euler--Lagrange equations
and yield the system \eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs} over the $n$th window for $t \in
[\timeStartArg{n},\timeEndArg{n}]$.
%\begin{align*}\
%&\massArg{n}  \genstateDotArg{n}{t}  -  \basisspaceTArg{n}
%\stweightingMatArgt{n}{t} \velocity(\veloargsromn) =  \massArg{n} \adjointArg{n}{t} , \\
% &\massArg{n} \adjointDotArg{n}{t}  + \basisspaceTArg{n} \bigg[\frac{\partial
%\velocity}{\partial \statey} (\veloargsromn) \bigg]^T \stweightingMatArgt{n}{t} \basisspaceArg{n}
% \adjointArg{n}{t} = \nonumber \\ 
%&-\basisspaceTArg{n} \big[
%\frac{\partial \velocity}{\partial \statey}(\veloargsromn) \big]^T \stweightingMatArgt{n}{t} \big( \mathbf{I} -
%\basisspaceArg{n} [\massArg{n}]^{-1} \basisspaceTArg{n} \stweightingMatArgt{n}{t} \big)
% \big( \basisspace \genstateDotArg{n}{t}  -
%\velocity(\veloargsromn) \big),  &
%%\genstateArg{n}{\timeStartArg{n}} = \begin{cases}
%%\genstate^{n-1}(\timeEndArg{n-1}) & n=2,\ldots,\nslabs, \\
%%\basisspace^T(\approxstate_0 - \stateIntercept) & n=1, \end{cases}\\ &
%%\adjointArg{n}{\timeEndArg{n}} = \boldsymbol 0.
%\end{align*} 
%subject to the boundary conditions~\eqref{eq:lspg_bcs1}--\eqref{eq:lspg_bcs}. 
\methodAcronym\ with \spatialAcronym\ trial subspaces can be alternatively interpreted as a Lagrange problem from optimal control: the Galerkin method is forced by a 
controller that enforces the residual minimization property. 
Section~\ref{sec:wls_spacetime} additionally considered \spaceTimeAcronym\
trial subspaces. For \spaceTimeAcronym\ trial subspaces, in where the
generalized coordinates are no longer functions, the stationary conditions
correspond to the system of algebraic equations \eqref{eq:st_stationary}.